---
title: "Salamander eye scaling and investment"
author: "Nicole Bednarik & Katie Thomas"
date: 8 August 2022
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: show
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(smatr)
library(picante)
library(nlme)
library(ape)
library(evobiR)
library(geiger)
library(caper)
library(phytools)
library(phylotools)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
library(ggtree)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(mvMORPH)
#library(evomap)
library(tidyverse)
library("dplyr")

# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)

```

# Data

First we import our tidied salamander morphology & trait dataset. 

```{r load-data}

# Import tidied salamander eye size and trait data
av.sal <- data.frame(read.csv("../Data/Tidy/salamanders_averages.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
# Quick look at data structure
str(av.sal)

```

Next, we import the amphibian tree from Jetz and Pyron (2019) that has been pruned to match the salamander species in this dataset.

```{r, fig.height=18, fig.width=7}

#Import pruned phylogeny
caudatatree <- read.nexus("../Data/Tidy/caudata-tree.nex")

#Plot tree
plot.phylo(caudatatree, show.tip.label=TRUE, cex = 0.7)

```

# Interspecific allometry using PGLS

In order to fit a PGLS regression in caper, we first need to make a comparative data object that includes our dataset and our phylogeny. Note that the gilled Pleurodeles waltl will be dropped from this, as there can only be one row for each species. 

```{r}
#Make row names of dataset the species names (so it will match phylogeny tips)
rownames(av.sal) <- av.sal$Genus_Species

#check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.sal)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp <- comparative.data(phy = caudatatree, data = av.sal, 
                            names.col = Genus_Species, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
sal.comp$dropped$Genus_Species#phylogeny
sal.comp$dropped$unmatched.rows #dataset

```

# Relative eye size and eye investment in salamanders

To investigate how relative eye size differed across species with different ecological traits, we looked at the residuals from PGLS fits of cornea diameter vs. body size. 

First, we added residuals of PGLS fits to our datasets by species.


```{r pgls-residuals}
#PGLS cornea diameter vs. cube root of mass ------------
pgls_cdmass <- pgls(log10(cor_av) ~ log10(rootmass_av), 
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#Residuals for PGLS of cornea diameter ~ cube root of mass

#extract pgls residuals 
pglsres.cdmass <- residuals(pgls_cdmass) 

#name residuals
colnames(pglsres.cdmass) <- "pglsres.cdmass" 

#Making PGLS vector into a datagrame
pglsres.cdmass <- as.data.frame(pglsres.cdmass)

#Make rownames into a column so we can use merge command
pglsres.cdmass <- tibble::rownames_to_column(pglsres.cdmass, "rn")

#Renaming Genus_Species by making a new column then deleting the old one
pglsres.cdmass$Genus_Species <- pglsres.cdmass$rn
pglsres.cdmass$rn <- NULL

#Merging datasets
pglsres.cdmass.all <- merge(av.sal, pglsres.cdmass, by = "Genus_Species")

# Export tidied dataset for analyses
write.csv(pglsres.cdmass.all, file = "../Data/Tidy/pglsres.cdmass.all.csv")

```

```{r}
#PGLS eye diameter vs. cube root of mass ------------
pgls_edmass <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#Residuals for PGLS of eye diameter ~ cube root of mass

#extract pgls residuals 
pglsres.edmass <- residuals(pgls_edmass) 

#name residuals
colnames(pglsres.edmass) <- "pglsres.edmass" 

#Making PGLS vector into a datagrame
pglsres.edmass <- as.data.frame(pglsres.edmass)

#Make rownames into a column so we can use merge command
pglsres.edmass <- tibble::rownames_to_column(pglsres.edmass, "rn")

#Renaming Genus_Species by making a new column then deleting the old one
pglsres.edmass$Genus_Species <- pglsres.edmass$rn
pglsres.edmass$rn <- NULL

#Merging datasets
pglsres.edmass.all <- merge(av.sal, pglsres.edmass, by = "Genus_Species")

# Export tidied dataset for analyses
write.csv(pglsres.edmass.all, file = "../Data/Tidy/pglsres.edmass.all.csv")
```

```{r}
#PGLS cornea diameter vs. SVL ------------
pgls_cdsvl <- pgls(log10(cor_av) ~ log10(svl_av), 
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#Residuals for PGLS of cornea diameter ~ snout-vent length

#extract pgls residuals 
pglsres.cdsvl <- residuals(pgls_cdsvl) 

#name residuals
colnames(pglsres.cdsvl) <- "pglsres.cdsvl" 

#Making PGLS vector into a datagrame
pglsres.cdsvl <- as.data.frame(pglsres.cdsvl)

#Make rownames into a column so we can use merge command
pglsres.cdsvl <- tibble::rownames_to_column(pglsres.cdsvl, "rn")

#Renaming Genus_Species by making a new column then deleting the old one
pglsres.cdsvl$Genus_Species <- pglsres.cdsvl$rn
pglsres.cdsvl$rn <- NULL

#Merging datasets
pglsres.cdsvl.all <- merge(av.sal, pglsres.cdsvl, by = "Genus_Species")

# Export tidied dataset for analyses
write.csv(pglsres.cdsvl.all, file = "../Data/Tidy/pglsres.cdsvl.all.csv")
```

```{r}
#PGLS eye diameter vs. SVL ------------
pgls_edsvl <- pgls(log10(eye_av) ~ log10(svl_av), 
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#Residuals for PGLS of eye diameter ~ snout-vent length

#extract pgls residuals 
pglsres.edsvl <- residuals(pgls_edsvl) 

#name residuals
colnames(pglsres.edsvl) <- "pglsres.edsvl" 

#Making PGLS vector into a datagrame
pglsres.edsvl <- as.data.frame(pglsres.edsvl)

#Make rownames into a column so we can use merge command
pglsres.edsvl <- tibble::rownames_to_column(pglsres.edsvl, "rn")

#Renaming Genus_Species by making a new column then deleting the old one
pglsres.edsvl$Genus_Species <- pglsres.edsvl$rn
pglsres.edsvl$rn <- NULL

#Merging datasets
pglsres.edsvl.all <- merge(av.sal, pglsres.edsvl, by = "Genus_Species")

# Export tidied dataset for analyses
write.csv(pglsres.edsvl.all, file = "../Data/Tidy/pglsres.edsvl.all.csv")
```

## Phylogenetic distribution of absolute and relative eye size

We then plotted absolute eye size and relative eye investment (represented by the residual of the eye size vs. body size PGLS) onto the salamander phylogeny. 

```{r plot-phytools, fig.height = 15, fig.width = 8}

# colorblind-friendly color vectors -------
# create vector of colors for adult habitat
col_hab <- c("Aquatic" = "#0072B2",
             "Scansorial" = "#009E73",
             "Semiaquatic" = "#56B4E9",
             "Subfossorial" = "#CC79A7",
             "Terrestrial" = "#E69F00")

#create vector of colors for activity period
col_act <- c("both" = "deeppink3",
             "diurnal" = "darkgoldenrod1", 
             "nocturnal" = "blueviolet")

# create vector of colors for larval habitat
col_larv <- c("Lentic water" = "royalblue1", 
         "Lotic water" = "skyblue1",
         "No larvae" = "lightcoral",
         "Obscured" = "darkslateblue",
         "On land" = "lightpink4")

# Prep data and phylogeny -----

#subset data for cornea diameter and mass
sal.bars.cdmass <- pglsres.cdmass.all %>%
  mutate(tip = Genus_Species, 
         abscor = cor_av, 
         invcor = pglsres.cdmass, 
         hab = Adult_Habitat,
         act = Activity_Period,
         larv = Larval_Habitat,
         family = Family) %>%
  select(tip, abscor, invcor, hab, act, larv, family)

# set row names in dataset to match the tip labels in the tree
row.names(sal.bars.cdmass) <- sal.bars.cdmass$tip

#drop phylogeny tips not in dataset
#tree_cdmass <- drop.tip(caudatatree, caudatatree$tip.label[!(caudatatree$tip.label %in% sal.bars.cdmass$tip)])

#ladderize tree
caudatatree <- ladderize(caudatatree)

#check that phylogeny and data match exactly
name.check(caudatatree, sal.bars.cdmass)

#resort trait dataset to the order of tree tip labels
sal.bars.cdmass <- sal.bars.cdmass[caudatatree$tip.label, ] 

#make trait vector for absolute cornea size
abscor <- as.vector(sal.bars.cdmass$abscor) 

#add tip label names to vector
names(abscor) <- sal.bars.cdmass$tip 

#make trait vector for cornea investment (PGLS residuals)
inv.mass <- as.vector(sal.bars.cdmass$invcor) #residuals of pgls
names(inv.mass) <- sal.bars.cdmass$tip

#make trait vector of habitats
habs.cdmass <- as.vector(sal.bars.cdmass$hab) 

#make vector of colors corresponding to phylogeny tips
tipcols.hab <- unname(col_hab[habs.cdmass]) 

# Make the phylogeny plot
p <- ggtree(caudatatree) %<+% labs + 
     geom_tiplab(size = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(375) + 
   coord_cartesian(clip = 'off') 


# Phylogeny with absolute cornea diameters -----

#plot tree with cornea size bars and tip labels
plotTree.wBars(caudatatree, abscor, 
               scale = 0.04, 
               tip.labels = TRUE, 
               offset = 0.35,
               ftype = "bi",
               fsize = 0.8,
               col = tipcols.hab)

#add legend for habitat states
legend(x = "left", legend = c("Aquatic", "Subfossorial", "Terrestrial", "Scansorial", "Semiaquatic"), pch = 22, pt.cex= 2, pt.bg = col_hab, cex = 1, bty = "n", horiz = F)

# Phylogeny with relative eye size (compared to mass) -------

#plot tree with relative eye size bars
plotTree.wBars(caudatatree, inv.mass, 
               scale = 0.7, 
               tip.labels = TRUE, 
               col = tipcols.hab,
               plot = TRUE, 
               add = TRUE)
```