---
title: "Plotting phylogenies by ecological state"
author: "Nicole Bednarik & Katie Thomas"
date: 28 May 2024
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: show
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(smatr)
library(picante)
library(nlme)
library(ape)
library(evobiR)
library(geiger)
library(caper)
library(phytools)
library(phylotools)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
library(ggtree)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(mvMORPH)
library(tidyverse)
library(dplyr)
library(ggplot2)

# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)

```

# Data

First we import our tidied salamander morphology & trait datasets with the residuals of PGLS fits. 

```{r load-data, results = "hide"}

# Import tidied data
pglsres.cdmass.all <- data.frame(read.csv("../Data/Tidy/pglsres.cdmass.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
pglsres.edmass.all <- data.frame(read.csv("../Data/Tidy/pglsres.edmass.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
pglsres.cdsvl.all <- data.frame(read.csv("../Data/Tidy/pglsres.cdsvl.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
pglsres.edsvl.all <- data.frame(read.csv("../Data/Tidy/pglsres.edsvl.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))

# Quick look at data structure
str(pglsres.cdmass.all)
str(pglsres.edmass.all)
str(pglsres.cdsvl.all)
str(pglsres.edsvl.all)

# Exponentiate PGLS residuals for plots and put in new dataframes
pglsres.cdmass.all.exp <- pglsres.cdmass.all %>%
  mutate(pglsres.cdmass = 10^pglsres.cdmass)
pglsres.edmass.all.exp <- pglsres.edmass.all %>%
  mutate(pglsres.edmass = 10^pglsres.edmass)
pglsres.cdsvl.all.exp <- pglsres.cdsvl.all %>%
  mutate(pglsres.cdsvl = 10^pglsres.cdsvl)
pglsres.edsvl.all.exp <- pglsres.edsvl.all %>%
  mutate(pglsres.edsvl = 10^pglsres.edsvl)
```

Here we add two rows to the dataframe to include the two blind salamander species (Proteus anguinus and Eurycea rathbuni) in order to eventually include them in the phylogeny for Figure 1.
```{r}

# Create new rows for blind salamanders
Eurycea_rathbuni <- data.frame(
  X.1 = 156, Genus_Species = "Eurycea_rathbuni", X = 159, eye_av = NA, cor_av = NA, 
  svl_av = 4.93, rootmass_av = 2.11, n = 3, Order = "Caudata", Suborder = "Salamandroidea", 
  Family = "Plethodontidae", Genus = "Eurycea", Species = "rathbuni", Gill_Presence = "yes", 
  Activity_Period = NA, Adult_Habitat = "aquatic", Development = "paedomorphic", 
  Life_History = "free-living", Larval_Habitat = "lentic water", pglsres.cdmass = NA
)

Proteus_anguinus <- data.frame(
  X.1 = 157, Genus_Species = "Proteus_anguinus", X = 160, eye_av = NA, cor_av = NA, 
  svl_av = 175.37, rootmass_av = 26.33, n = 3, Order = "Caudata", Suborder = "Salamandroidea", 
  Family = "Proteidae", Genus = "Proteus", Species = "anguinus", Gill_Presence = "yes", 
  Activity_Period = "nocturnal", Adult_Habitat = "aquatic", Development = "paedomorphic", 
  Life_History = "free-living", Larval_Habitat = "lentic water", pglsres.cdmass = NA
)

# Add the new rows to the existing dataframe
pglsres.cdmass.all.exp.blind <- rbind(pglsres.cdmass.all.exp, Eurycea_rathbuni, Proteus_anguinus)
```

Next, we import two amphibian trees from Jetz and Pyron (2019) that have been pruned to match the salamander species in this dataset. One includes the two blind salamander species (Proteus anguinus and Eurycea rathbuni) and will be used to create Figure 1 for the paper.

```{r, fig.height=18, fig.width=7}

# Import pruned phylogeny
caudatatree <- read.nexus("../Data/Tidy/caudata-tree.nex")
# Import prned phylogeny (including blind salamanders)
caudatatree_blind <- read.nexus("../Data/Tidy/caudata-tree-blind.nex")

# Plot tree
plot.phylo(caudatatree, show.tip.label=TRUE, cex = 0.7)
plot.phylo(caudatatree_blind, show.tip.label=TRUE, cex = 0.7)

```

# Phylogenetic distribution of absolute and relative eye size

We then plot absolute eye size and relative eye investment (represented by the residual of the eye size vs. body size PGLS) onto the salamander phylogeny. 

## Color figures with pallettes in the frog paper

We start by running a custom color palette for the 4 ecological states to be used for all following figures. 

First we define what we want those color palettes to be:

```{r color-pallette, results = "hide"}

# Rename metamorphosizing to metamorphosing
pglsres.cdmass.all.exp <- pglsres.cdmass.all.exp %>%
  mutate(Development = case_when(
    Development == "metamorphosizing" ~ "metamorphosing",
    TRUE ~ Development
  ))
# Rename no larva to no larvae
pglsres.cdmass.all.exp <- pglsres.cdmass.all.exp %>%
  mutate(Larval_Habitat = case_when(
    Larval_Habitat == "no larva" ~ "no larvae",
    TRUE ~ Larval_Habitat
  ))

# Define a colorblind-friendly vector of colors for adult habitat, activity period, larval habitat, and development mode
col_hab <- c("aquatic" = "#0072B2",
             "scansorial" = "#009E73",
             "semiaquatic" = "#56B4E9",
             "subfossorial" = "#CC79A7",
             "terrestrial" = "#E69F00")
col_act <- c("both" = "deeppink3",
             "diurnal" = "darkgoldenrod1", 
             "nocturnal" = "blueviolet",
             "NA" = "grey")
col_larv <- c("lentic water" = "royalblue1", 
         "lotic water" = "skyblue1",
         "no larvae" = "lightcoral",
         "both" = "forestgreen")
col_dev <- c("metamorphosing" = "darkturquoise",
             "paedomorphic" = "black",
             "direct development" = "#332288")
```

## Phylogenetic distribution of eye size by adult habitat

Here we create the salamander phylogeny with absolute cornea size and relative cornea investment (mass) for Figure 1, colored by Adult Habitat. We first need to make a dataframe that includes the phylogeny.

```{r sal-phy, results = "hide"}

# Make dataframe for labeling phylogeny figure
sal.phy <- pglsres.cdmass.all.exp %>%
  # Make column labeling tip labels that match phylogeny
  mutate(tip = Genus_Species) %>%
  separate(Genus_Species, c("genus", "species"), sep = "_", extra = "drop") %>%
  # Add tip labels with regular text
  mutate(labels = as.factor(paste(genus, species, sep = " "))) %>%
  select(tip, labels, cor_av, pglsres.cdmass, Adult_Habitat, Activity_Period, Larval_Habitat, Development) 
sal.phy.blind <- pglsres.cdmass.all.exp.blind %>%
  # Make column labeling tip labels that match phylogeny
  mutate(tip = Genus_Species) %>%
  separate(Genus_Species, c("genus", "species"), sep = "_", extra = "drop") %>%
  # Add tip labels with regular text
  mutate(labels = as.factor(paste(genus, species, sep = " "))) %>%
  select(tip, labels, cor_av, pglsres.cdmass, Adult_Habitat, Activity_Period, Larval_Habitat, Development) 

# Set row names in dataset to match the tip labels in the tree
row.names(sal.phy) <- sal.phy$tip
row.names(sal.phy.blind) <- sal.phy.blind$tip

# Check that phylogeny and data match exactly
name.check(caudatatree, sal.phy)
name.check(caudatatree_blind, sal.phy.blind)
```

Next, we use ggtree to plot the phylogeny alongside bar plots for absolute cornea diameter and relative cornea size and color by adult habitat.  

```{r, fig.height=18, fig.width=6}

# Make the phylogeny plot
p <- ggtree(caudatatree) %<+% labs + 
     geom_tiplab(size = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(375) + 
   coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (bars for cornea diameter)
p2 <- facet_plot(p, panel="Cornea diameter (mm)", data=sal.phy, geom=geom_segment, aes(x=0, xend=cor_av, y=y, yend=y, color = Adult_Habitat), size = 3, alpha = 0.7) 

# Make a third plot next to the first two (bars for relative cornea size)
p3 <- facet_plot(p2, panel='Relative cornea size', data=sal.phy, geom=geom_segment, aes(x=1, xend=pglsres.cdmass, y=y, yend=y, color = Adult_Habitat), size=3, alpha = 0.7) +
  scale_color_manual(values = col_hab) +
  theme_tree2(legend.position=c(.07, .93)) 

# Control widths of panels
 gt = ggplot_gtable(ggplot_build(p3))
 
#gtable_show_layout(gt) # Will show you the layout - very handy function
#gt # See plot layout in table format
 gt$layout$l[grep('tree', gt$layout$name)] # You want to find the column specific to panel-2
 gt$widths[7] = 0.5*gt$widths[5] # Reduce column 7
 gt$widths[9] = 0.5*gt$widths[5]
 
# Print figure
plot(gt)
# Export plot to pdf 
ggsave("../Plots/hab-tree-figure.pdf", plot = gt, device = "pdf", width = 12, height = 18)
```

We copy the same code here but use the phylogeny including the two blind salamander species, which will be used to create Figure 1 for the paper. 

```{r, fig.height=18, fig.width=6}

# Make the phylogeny plot
p_blind <- ggtree(caudatatree_blind) %<+% labs + 
     geom_tiplab(size = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(375) + 
   coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (bars for cornea diameter)
p2_blind <- facet_plot(p_blind, panel="Cornea diameter (mm)", data=sal.phy.blind, geom=geom_segment, aes(x=0, xend=cor_av, y=y, yend=y, color = Adult_Habitat), size = 3, alpha = 0.7) 

# Make a third plot next to the first two (bars for relative cornea size)
p3_blind <- facet_plot(p2_blind, panel='Relative cornea size', data=sal.phy.blind, geom=geom_segment, aes(x=1, xend=pglsres.cdmass, y=y, yend=y, color = Adult_Habitat), size=3, alpha = 0.7) +
  scale_color_manual(values = col_hab) +
  theme_tree2(legend.position = "none") +
  theme(
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )

# Control widths of panels
 gt_blind = ggplot_gtable(ggplot_build(p3_blind))
 
#gtable_show_layout(gt_blind) # Will show you the layout - very handy function
#gt_blind # See plot layout in table format
 gt_blind$layout$l[grep('tree', gt_blind$layout$name)] # You want to find the column specific to panel-2
 gt_blind$widths[7] = 0.5*gt$widths[5] # Reduce column 7
 gt_blind$widths[9] = 0.5*gt$widths[5]
 
# Print figure
plot(gt_blind)
# Export plot to pdf 
ggsave("../Plots/hab-tree-blind-figure.pdf", plot = gt_blind, device = "pdf", width = 12, height = 18)
```

We repeat the same steps for phylogenies colored by activity period, larval habitat, and development.

## Phylogenetic distribution of eye size by activity period

```{r}
####TEST

# Create a base plot to get y positions
p <- ggtree(caudatatree)

# Extract data from the plot
tree_data <- p$data

sal.phy <- sal.phy %>%
  rename(label = tip)

# Merge tree data with sal.phy based on the tip labels
sal.phy <- sal.phy %>%
  left_join(tree_data %>% select(label, y), by = "label")
head(sal.phy)

# Make the phylogeny plot
p_act <- ggtree(caudatatree) %<+% labs + 
  geom_tiplab(size = 3, aes(label = labels), fontface = 3) + 
  xlim_tree(375) + 
  coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (bars for cornea diameter)
p2_act <- facet_plot(p_act, panel="Cornea diameter (mm)", data=sal.phy, 
                     geom=geom_segment, 
                     aes(x=0, xend=cor_av, y=y, yend=y, color = Activity_Period), 
                     size = 3, alpha = 0.7) +
                     geom_segment(data = sal.phy,
                      aes(x = 0, xend = cor_av, y = y, yend = y),
                      color = "black", size = 0.5)

# Make a third plot next to the first two (bars for relative cornea size)
p3_act <- facet_plot(p2_act, panel='Relative cornea size', data=sal.phy, 
                     geom=geom_segment, 
                     aes(x=1, xend=pglsres.cdmass, y=y, yend=y, color = Activity_Period), 
                     size=3, alpha = 0.7) +
          geom_segment(data=sal.phy, 
                       aes(x=1, xend=pglsres.cdmass, y=y, yend=y), size=0.5, alpha=1) +
            geom_segment(data = sal.phy,
               aes(x = 1, xend = pglsres.cdmass, y = y, yend = y),
               color = "black", size = 0.5) +
  scale_color_manual(values = col_act) +
  theme_tree2(legend.position = "none") +
  theme(
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  ) 

# Control widths of panels
gt_act = ggplot_gtable(ggplot_build(p3_act))

#gtable_show_layout(gt_act) # Will show you the layout - very handy function
#gt_act # See plot layout in table format
gt_act$layout$l[grep('tree', gt_act$layout$name)] # You want to find the column specific to panel-2
gt_act$widths[7] = 0.5*gt_act$widths[5] # Reduce column 7
gt_act$widths[9] = 0.5*gt_act$widths[5]

# Print figure
plot(gt_act)

# Export plot to pdf 
ggsave("../Plots/act-tree-figure.pdf", plot = gt_act, device = "pdf", width = 12, height = 18)

####TEST
```

```{r, fig.height=18, fig.width=6}
# Make the phylogeny plot
p_act <- ggtree(caudatatree) %<+% labs + 
     geom_tiplab(size = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(375) + 
   coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (bars for cornea diameter)
p2_act <- facet_plot(p_act, panel="Cornea diameter (mm)", data=sal.phy, geom=geom_segment, aes(x=0, xend=cor_av, y=y, yend=y, color = Activity_Period), size = 3, alpha = 0.7) 
plot(p2_act)
print(sal.phy$y)

# Make a third plot next to the first two (bars for relative cornea size)
p3_act <- facet_plot(p2_act, panel='Relative cornea size', data=sal.phy, geom=geom_segment, aes(x=1, xend=pglsres.cdmass, y=y, yend=y, color = Activity_Period), size=3, alpha = 0.7) +
  scale_color_manual(values = col_act) +
  theme_tree2(legend.position = "none") +
  theme(
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  ) 

# Control widths of panels
 gt_act = ggplot_gtable(ggplot_build(p3_act))
 
#gtable_show_layout(gt_act) # Will show you the layout - very handy function
#gt_act # See plot layout in table format
 gt_act$layout$l[grep('tree', gt_act$layout$name)] # You want to find the column specific to panel-2
 gt_act$widths[7] = 0.5*gt_act$widths[5] # Reduce column 7
 gt_act$widths[9] = 0.5*gt_act$widths[5]
 
# Print figure
plot(gt_act)
# Export plot to pdf 
ggsave("../Plots/act-tree-figure.pdf", plot = gt_act, device = "pdf", width = 12, height = 18)
```

## Phylogenetic distribution of eye size by larval habitat

```{r, fig.height=18, fig.width=6}

# Filter out species with NA for larval habitat
sal.phy.larv <- sal.phy %>% filter(!is.na(Larval_Habitat))

# Make the phylogeny plot
p_larv <- ggtree(caudatatree) %<+% labs + 
     geom_tiplab(size = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(375) + 
   coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (bars for cornea diameter)
p2_larv <- facet_plot(p_larv, panel="Cornea diameter (mm)", data=sal.phy.larv, geom=geom_segment, aes(x=0, xend=cor_av, y=y, yend=y, color = Larval_Habitat), size = 3, alpha = 0.7) 

# Make a third plot next to the first two (bars for relative cornea size)
p3_larv <- facet_plot(p2_larv, panel='Relative cornea size', data=sal.phy.larv, geom=geom_segment, aes(x=1, xend=pglsres.cdmass, y=y, yend=y, color = Larval_Habitat), size=3, alpha = 0.7) +
  scale_color_manual(values = col_larv)  +
  theme_tree2(legend.position = "none") +
  theme(
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )

# Control widths of panels
 gt_larv = ggplot_gtable(ggplot_build(p3_larv))
 
#gtable_show_layout(gt_larv) # Will show you the layout - very handy function
#gt_larv # See plot layout in table format
 gt_larv$layout$l[grep('tree', gt_larv$layout$name)] # You want to find the column specific to panel-2
 gt_larv$widths[7] = 0.5*gt_larv$widths[5] # Reduce column 7
 gt_larv$widths[9] = 0.5*gt_larv$widths[5]
 
# Print figure
plot(gt_larv)
# Export plot to pdf 
ggsave("../Plots/larv-tree-figure.pdf", plot = gt_larv, device = "pdf", width = 12, height = 18)
```

## Phylogenetic distribution of eye size by development mode

```{r, fig.height=18, fig.width=6}

# Filter out species with NA for development mode
sal.phy.dev <- sal.phy %>% filter(!is.na(Development))

# Make the phylogeny plot
p_dev <- ggtree(caudatatree) %<+% labs + 
     geom_tiplab(size = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(375) + 
   coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (bars for cornea diameter)
p2_dev <- facet_plot(p_dev, panel="Cornea diameter (mm)", data=sal.phy.dev, geom=geom_segment, aes(x=0, xend=cor_av, y=y, yend=y, color = Development), size = 3, alpha = 0.7) 

# Make a third plot next to the first two (bars for relative cornea size)
p3_dev <- facet_plot(p2_dev, panel='Relative cornea size', data=sal.phy.dev, geom=geom_segment, aes(x=1, xend=pglsres.cdmass, y=y, yend=y, color = Development), size=3, alpha = 0.7) +
  scale_color_manual(values = col_dev)  +
  theme_tree2(legend.position = "none") +
  theme(
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  ) 

# Control widths of panels
 gt_dev = ggplot_gtable(ggplot_build(p3_dev))
 
#gtable_show_layout(gt_dev) # Will show you the layout - very handy function
#gt_dev # See plot layout in table format
 gt_dev$layout$l[grep('tree', gt_dev$layout$name)] # You want to find the column specific to panel-2
 gt_dev$widths[7] = 0.5*gt_dev$widths[5] # Reduce column 7
 gt_dev$widths[9] = 0.5*gt_dev$widths[5]
 
# Print figure
plot(gt_dev)
# Export plot to pdf 
ggsave("../Plots/dev-tree-figure.pdf", plot = gt_dev, device = "pdf", width = 12, height = 18)
```