---
title: "Boxplots"
author: "Nicole Bednarik & Katie Thomas"
date: 9 September 2022
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: show
---

```{r setup, include = FALSE}
# Load package libraries
library(plyr)
library(geiger)
library(caper)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
library(ggtree)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(tidyverse)
library(plotly)
# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)
```

# Data

First we import our tidied salamander morphology & trait binary dataset with the residuals of a cornea ~ cube root of mass PGLS as a column.

```{r load-data}
# Import tidied salamander eye siz, investment, and trait data
pglsres.cdmass.all <- data.frame(read.csv("../Data/Tidy/pglsres.cdmass.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
pglsres.edmass.all <- data.frame(read.csv("../Data/Tidy/pglsres.edmass.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
pglsres.cdsvl.all <- data.frame(read.csv("../Data/Tidy/pglsres.cdsvl.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
pglsres.edsvl.all <- data.frame(read.csv("../Data/Tidy/pglsres.edsvl.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
# Quick look at data structure
str(pglsres.cdmass.all)
str(pglsres.edmass.all)
str(pglsres.cdsvl.all)
str(pglsres.edsvl.all)
```

Next, we make boxplots of how eye size and relative eye size (represented by the PGLS residuals) compare across salamander ecological traits.

```{r trait-boxplots}

# colorblind-friendly color vectors -------
# create vector of colors for adult habitat
col_hab <- c("aquatic" = "#0072B2",
             "scansorial" = "#009E73",
             "semiaquatic" = "#56B4E9",
             "subfossorial" = "#CC79A7",
             "terrestrial" = "#E69F00")

#create vector of colors for activity period
col_act <- c("both" = "deeppink3",
             "diurnal" = "darkgoldenrod1", 
             "nocturnal" = "blueviolet")

# create vector of colors for larval habitat
col_larv <- c("lentic water" = "royalblue1", 
         "lotic water" = "skyblue1",
         "no larva" = "lightcoral",
         "both" = "darkslateblue")

# create vector of colors for metamorphosing/paedomorphic aquatic species
col_dev <- c("metamorphosizing" = "darkturquoise", 
         "paedomorphic" = "black")

```

# Binary Habitat States

We recategorized our species data into binary habitat categories used by Huang et al, a study of frog corneal eye size. We looked at if eye size and investment differed across these categories.

```{r}
# Group habitat type into binary categories used by Huang et al.
pglsres.cdmass.all$Habitat_Group <- ifelse(pglsres.cdmass.all$Adult_Habitat %in% c("aquatic", "semiaquatic"), "Aq/S", "T/Ar")
# Group habitat type into binary categories used by Huang et al.
pglsres.cdsvl.all$Habitat_Group <- ifelse(pglsres.cdsvl.all$Adult_Habitat %in% c("aquatic", "semiaquatic"), "Aq/S", "T/Ar")
# Group habitat type into binary categories used by Huang et al.
pglsres.edmass.all$Habitat_Group <- ifelse(pglsres.edmass.all$Adult_Habitat %in% c("aquatic", "semiaquatic"), "Aq/S", "T/Ar")
# Group habitat type into binary categories used by Huang et al.
pglsres.edsvl.all$Habitat_Group <- ifelse(pglsres.edsvl.all$Adult_Habitat %in% c("aquatic", "semiaquatic"), "Aq/S", "T/Ar")

# boxplot of cornea size vs adult habitat
abscor.habitat <- ggplot(data=pglsres.cdmass.all, aes(x = Habitat_Group, y = cor_av)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, colour="black", geom="point", 
               shape=18, size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Adult_Habitat), 
              shape = 19, size = 1.3, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Cornea diameter (mm)") +
  xlab("Binary habitat") +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  theme(legend.position = "none")
#print plot
abscor.habitat

pglsres.cdmass.all$Habitat_Group <- ifelse(pglsres.cdmass.all$Adult_Habitat %in% c("aquatic", "semiaquatic"), "Aq/S", "T/Ar")

# boxplot of eye size vs adult habitat
abseye.habitat <- ggplot(data=pglsres.cdmass.all, aes(x = Habitat_Group, y = eye_av)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, colour="black", geom="point", 
               shape=18, size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Adult_Habitat), 
              shape = 19, size = 1.3, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Eye diameter (mm)") +
  xlab("Binary habitat") +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  theme(legend.position = "none")
#print plot
abseye.habitat

# boxplot of cornea investment by mass vs adult habitat
cdmass.habitat <- ggplot(data=pglsres.cdmass.all, aes(x = Habitat_Group, y = pglsres.cdmass)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, colour="black", geom="point", 
               shape=18, size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Adult_Habitat), 
              shape = 19, size = 1.3, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Cornea investment (mass)") +
  xlab("Binary habitat") +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  theme(legend.position = "none")
#print plot
cdmass.habitat

# boxplot of cornea investment by svl vs adult habitat
cdsvl.habitat <- ggplot(data=pglsres.cdsvl.all, aes(x = Habitat_Group, y = pglsres.cdsvl)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, colour="black", geom="point", 
               shape=18, size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Adult_Habitat), 
              shape = 19, size = 1.3, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Cornea investment (SVL)") +
  xlab("Binary habitat") +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  theme(legend.position = "none")
#print plot
cdsvl.habitat

# boxplot of eye investment by mass vs adult habitat
edmass.habitat <- ggplot(data=pglsres.edmass.all, aes(x = Habitat_Group, y = pglsres.edmass)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, colour="black", geom="point", 
               shape=18, size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Adult_Habitat), 
              shape = 19, size = 1.3, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Eye investment (mass)") +
  xlab("Binary habitat") +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  theme(legend.position = "none")
#print plot
edmass.habitat

# boxplot of eye investment by svl vs adult habitat
edsvl.habitat <- ggplot(data=pglsres.edsvl.all, aes(x = Habitat_Group, y = pglsres.edsvl)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, colour="black", geom="point", 
               shape=18, size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Adult_Habitat), 
              shape = 19, size = 1.3, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Eye investment (SVL)") +
  xlab("Binary habitat") +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  theme(legend.position = "none")
#print plot
edsvl.habitat

# Export plot to pdf - you can change the export dimensions by changing the height and width numbers to make it fit well on your poster!
ggsave("../Plots/binary.abscor.boxplot.pdf", plot = abscor.habitat, device = "pdf")
ggsave("../Plots/binary.abseye.boxplot.pdf", plot = abseye.habitat, device = "pdf")
ggsave("../Plots/binary.cdmass.boxplot.pdf", plot = cdmass.habitat, device = "pdf")
ggsave("../Plots/binary.cdsvl.boxplot.pdf", plot = cdsvl.habitat, device = "pdf")
ggsave("../Plots/binary.edmass.boxplot.pdf", plot = edmass.habitat, device = "pdf")
ggsave("../Plots/binary.edsvl.boxplot.pdf", plot = edsvl.habitat, device = "pdf")

```

Next we will create boxplots comparing eye and cornea size/investment and the ecological traits. 

##Cornea size and investment across adult habitats
```{r}
# Boxplot of cornea diameter vs adult habitat
adult_habitat.abs <- ggplot(data = filter(pglsres.cdmass.all, !is.na(Adult_Habitat)), 
                             aes(x = reorder(Adult_Habitat, cor_av, FUN = mean), y = cor_av)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, shape = Adult_Habitat, color = Adult_Habitat), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  ylab("Cornea diameter (mm)") +
  xlab("Adult habitat") +
  theme(legend.position = "none")

# use our custom shapes for points
  #scale_shape_manual(values = shape_hab, name = "Adult habitat")

# Boxplot of cornea investment (by mass) vs adult habitat
adult_habitat.inv <- ggplot(data = filter(pglsres.cdmass.all, !is.na(Adult_Habitat)), 
                             aes(x = reorder(Adult_Habitat, pglsres.cdmass, FUN = mean), y = pglsres.cdmass)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Adult_Habitat), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  ylab("Cornea investment (mass)") +
  xlab("Adult habitat") +
  theme(legend.position = "none")

# Boxplot of cornea investment (by SVL) vs adult habitat
adult_habitat.invsvl <- ggplot(data = filter(pglsres.cdsvl.all, !is.na(Adult_Habitat)), 
                             aes(x = reorder(Adult_Habitat, pglsres.cdsvl, FUN = mean), y = pglsres.cdsvl)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Adult_Habitat), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_hab) +  # Use the color vector for manual color coloring based on Adult_Habitat
  ylab("Cornea investment (SVL)") +
  xlab("Adult Habitat") +
  theme(legend.position = "none")

# Look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
ggplotly(adult_habitat.abs)
ggplotly(adult_habitat.inv)
ggplotly(adult_habitat.invsvl)

# Export plot to pdf - you can change the export dimensions by changing the height and width numbers to make it fit well on your poster!
ggsave("../Plots/adulthabitat.abs.boxplot.pdf", plot = adult_habitat.abs, device = "pdf")
ggsave("../Plots/adulthabitat.invmass.boxplot.pdf", plot = adult_habitat.inv, device = "pdf")
ggsave("../Plots/adulthabitat.invsvl.boxplot.pdf", plot = adult_habitat.invsvl, device = "pdf")
```

##Cornea size and investment by activity period

```{r}
# Boxplot of cornea diameter vs activity period
activity_period.abs <- ggplot(data = filter(pglsres.cdmass.all, !is.na(Activity_Period)), 
                             aes(x = reorder(Activity_Period, cor_av, FUN = mean), y = cor_av)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, shape = Activity_Period, color = Activity_Period), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_act) +  # Use the color vector for manual color coloring based on Activity Period
  ylab("Cornea diameter (mm)") +
  xlab("Activity period") +
  theme(legend.position = "none")

# Boxplot of cornea investment (by mass) vs activity period
activity_period.inv <- ggplot(data = filter(pglsres.cdmass.all, !is.na(Activity_Period)), 
                             aes(x = reorder(Activity_Period, pglsres.cdmass, FUN = mean), y = pglsres.cdmass)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Activity_Period), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_act) +  # Use the color vector for manual color coloring based on Activity Period
  ylab("Cornea investment (mass)") +
  xlab("Activity period") +
  theme(legend.position = "none")

# Boxplot of cornea investment (by SVL) vs activity period
activity_period.invsvl <- ggplot(data = filter(pglsres.cdsvl.all, !is.na(Activity_Period)), 
                             aes(x = reorder(Activity_Period, pglsres.cdsvl, FUN = mean), y = pglsres.cdsvl)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Activity_Period), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_act) +  # Use the color vector for manual color coloring based on Activity Period
  ylab("Cornea investment (SVL)") +
  xlab("Activity period") +
  theme(legend.position = "none")

# Look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
ggplotly(activity_period.abs)
ggplotly(activity_period.inv)
ggplotly(activity_period.invsvl)

# Export plot to pdf - you can change the export dimensions by changing the height and width numbers to make it fit well on your poster!
ggsave("../Plots/activityperiod.abs.boxplot.pdf", plot = activity_period.abs, device = "pdf")
ggsave("../Plots/activityperiod.invmass.boxplot.pdf", plot = activity_period.inv, device = "pdf")
ggsave("../Plots/activityperiod.invsvl.boxplot.pdf", plot = activity_period.invsvl, device = "pdf")
```

##Cornea size and investment by larval habitat

```{r}
# Boxplot of cornea diameter vs larval habitat
larval_habitat.abs <- ggplot(data = filter(pglsres.cdmass.all, !is.na(Larval_Habitat)), 
                             aes(x = reorder(Larval_Habitat, cor_av, FUN = mean), y = cor_av)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, shape = Larval_Habitat, color = Larval_Habitat), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_larv) +  # Use the color vector for manual color coloring based on Larval Habitat
  ylab("Cornea diameter (mm)") +
  xlab("Larval habitat") +
  theme(legend.position = "none")

# Boxplot of cornea investment (by mass) vs larval habitat
larval_habitat.inv <- ggplot(data = filter(pglsres.cdmass.all, !is.na(Larval_Habitat)), 
                             aes(x = reorder(Larval_Habitat, pglsres.cdmass, FUN = mean), y = pglsres.cdmass)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Larval_Habitat), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_larv) +  # Use the color vector for manual color coloring based on Larval Habitat
  ylab("Cornea investment (mass)") +
  xlab("Larval habitat") +
  theme(legend.position = "none")

# Boxplot of cornea investment (by SVL) vs larval habitat
larval_habitat.invsvl <- ggplot(data = filter(pglsres.cdsvl.all, !is.na(Larval_Habitat)), 
                             aes(x = reorder(Larval_Habitat, pglsres.cdsvl, FUN = mean), y = pglsres.cdsvl)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Larval_Habitat), 
              shape = 19, size = 2, alpha = 0.5, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_larv) +  # Use the color vector for manual color coloring based on Larval Habitat
  ylab("Cornea investment (SVL)") +
  xlab("Larval habitat") +
  theme(legend.position = "none")

# Look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
ggplotly(larval_habitat.abs)
ggplotly(larval_habitat.inv)
ggplotly(larval_habitat.invsvl)

# Export plot to pdf - you can change the export dimensions by changing the height and width numbers to make it fit well on your poster!
ggsave("../Plots/larvalhabitat.abs.boxplot.pdf", plot = larval_habitat.abs, device = "pdf")
ggsave("../Plots/larvalhabitat.invmass.boxplot.pdf", plot = larval_habitat.inv, device = "pdf")
ggsave("../Plots/larvalhabitat.invsvl.boxplot.pdf", plot = larval_habitat.invsvl, device = "pdf")

```


##Cornea size and investment by development status of aquatic species

```{r}
# Filtering out only aquatic species into a new dataset
aquatic_species <- pglsres.cdmass.all %>%
  filter(Adult_Habitat == "aquatic")
aquatic_species.svl <- pglsres.cdsvl.all %>%
  filter(Adult_Habitat == "aquatic")

# Boxplot of absolute cornea size vs. paedomorphic/metamorphosizing status
aquaticdevelopment.abs <- ggplot(data = aquatic_species, aes(x = reorder(Development, cor_av, FUN = mean), y = cor_av)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Development), 
              shape = 19, size = 2, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_dev) +  # Use the color vector for manual color coloring based on Adult_Habitat
  ylab("Cornea investment (mass)") +
  xlab("Development Type") +
  theme(legend.position = "none")

# Boxplot of pgls residuals by mass vs. paedomorphic/metamorphosizing status
aquaticdevelopment.mass <- ggplot(data = aquatic_species, aes(x = reorder(Development, pglsres.cdmass, FUN = mean), y = pglsres.cdmass)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Development), 
              shape = 19, size = 2, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_dev) +  # Use the color vector for manual color coloring based on Adult_Habitat
  ylab("Cornea investment (mass)") +
  xlab("Development Type") +
  theme(legend.position = "none")

# Boxplot of pgls residuals by svl vs. paedomorphic/metamorphosizing status
aquaticdevelopment.svl <- ggplot(data = aquatic_species.svl, aes(x = reorder(Development, pglsres.cdsvl, FUN = mean), y = pglsres.cdsvl)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun.y = mean, color = "black", geom = "point", 
               shape = 18, size = 3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, color = Development), 
              shape = 19, size = 2, alpha = 0.7, position = position_jitter(0.15)) +
  theme(text = element_text(size = 14), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_color_manual(values = col_dev) +  # Use the color vector for manual color coloring based on Adult_Habitat
  ylab("Cornea investment (SVL)") +
  xlab("Development Type") +
  theme(legend.position = "none")

# Look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
ggplotly(aquaticdevelopment.abs)
ggplotly(aquaticdevelopment.mass)
ggplotly(aquaticdevelopment.svl)

# Export plot to pdf - you can change the export dimensions by changing the height and width numbers to make it fit well on your poster!
ggsave("../Plots/aquaticdevelopment.abs.boxplot.pdf", plot = aquaticdevelopment, device = "pdf")
ggsave("../Plots/aquaticdevelopment.mass.boxplot.pdf", plot = aquaticdevelopment, device = "pdf")
ggsave("../Plots/aquaticdevelopment.svl.boxplot.pdf", plot = aquaticdevelopment.svl, device = "pdf")
```

Next, we use ggtree to plot the phylogeny alongside bar plots for absolute cornea diameter and relative conea size. 

```{r, fig.height=18, fig.width=6}
library(ggtree)
# Make the phylogeny plot
p <- ggtree(caudatatree) %<+% labs + 
     geom_tiplab(size = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(375) + 
   coord_cartesian(clip = 'off') 
# Make a second plot next to the phylogeny (bars for cornea diameter)
p2 <- facet_plot(p, panel="Cornea diameter (mm)", data=sal.phy, geom=geom_segment, aes(x=0, xend=cor_av, y=y, yend=y, color = hab), size = 3, alpha = 0.7) +
  scale_color_manual(values = col_hab,
                     name = "Adult habitat")
# Make a third plot next to the first two (bars for relative cornea size)
p3 <- facet_plot(p2, panel='Relative cornea size', data=sal.phy, geom=geom_segment, aes(x=0, xend=pglsres.cdmass, y=y, yend=y, color = hab), size=3, alpha = 0.7) +
  theme_tree2(legend.position=c(.07, .93)) +# add scale bars and move legends
  scale_color_manual(values = col_hab,
                     name = "Adult habitat")
#control widths of panels
 gt = ggplot_gtable(ggplot_build(p3))
# gtable_show_layout(gt) # will show you the layout - very handy function
# gt # see plot layout in table format
# gt$layout$l[grep('tree', gt$layout$name)] # you want to find the column specific to panel-2
 gt$widths[7] = 0.5*gt$widths[5] # reduce column 7
 gt$widths[9] = 0.5*gt$widths[5]
#print figure
plot(gt)
#export figure
pdf("tree-figure.pdf", width = 8, height = 22)
grid::grid.draw(gt) # plot with grid draw
dev.off()
```

