---
title: "PGLS Analysis"
author: "Nicole Bednarik & Katie Thomas"
date: 28 May 2024
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: show
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(geiger)
library(caper)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(tidyverse)
library(car)
library(multcomp)
install.packages("dunn.test")
library(dunn.test)

# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)
```

# Data

First we import our tidied salamander morphology & trait dataset with the residuals of a cornea ~ cube root of mass PGLS as a column.

```{r load-data}

# Import tidied salamander eye size and trait data
pglsres.cdmass.all <- data.frame(read.csv("../Data/Tidy/pglsres.cdmass.all.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))

# Import pruned phylogeny
caudatatree <- read.nexus("../Data/Tidy/caudata-tree.nex")

# Quick look at data structure
str(pglsres.cdmass.all)

# Filtering out only aquatic species into a new dataset
aquatic_species <- pglsres.cdmass.all %>%
  filter(Adult_Habitat == "aquatic")
```

We combine the phylogeny and tidied dataset into one comparative object to be used in ensuing PGLS analyses. This is also done separately for only aquatic species.

```{r sal-comp, results = "hide"}

# Make row names of dataset the species names (so it will match phylogeny tips)
rownames(pglsres.cdmass.all) <- pglsres.cdmass.all$Genus_Species

# Check that names match in dataframe and tree
name.check(phy = caudatatree, data = pglsres.cdmass.all)

# Reorder factor levels
pglsres.cdmass.all$Activity_Period <- factor(pglsres.cdmass.all$Activity_Period, levels = c("nocturnal", "diurnal", "both"))
pglsres.cdmass.all$Larval_Habitat <- factor(pglsres.cdmass.all$Larval_Habitat, levels = c("lentic water", "lotic water", "both", "no larva"))

# Use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp <- comparative.data(phy = caudatatree, data = pglsres.cdmass.all, names.col = Genus_Species, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)
aquatic.comp <- comparative.data(phy = caudatatree, data = aquatic_species, names.col = Genus_Species, vcv = TRUE, na.omit = FALSE, warn.dropped = TRUE)

# Check for dropped tips or dropped species
sal.comp$dropped$tips 
aquatic.comp$dropped$tips
```

## Testing for differences in eye size across ecological states

### Assumptions

First we check if data for eye size and eye investment across ecological states is normally distributed and that there are equal variances across samples. 

```{r}

# Testing Assumptions

# EYE SIZE
# Create a QQ plot 
qqnorm(pglsres.cdmass.all$eye_av)
qqline(pglsres.cdmass.all$eye_av)
# Histogram 
hist(pglsres.cdmass.all$eye_av, main = "Histogram of Eye Size")
# Check the normality 
shapiro.test(pglsres.cdmass.all$eye_av)
# Check equality of variances
leveneTest(eye_av ~ Adult_Habitat, data = pglsres.cdmass.all)

#CORNEA SIZE
# Create a QQ plot 
qqnorm(pglsres.cdmass.all$cor_av)
qqline(pglsres.cdmass.all$cor_av)
# Histogram 
hist(pglsres.cdmass.all$cor_av, main = "Histogram of Cornea Size")
# Check the normality
shapiro.test(pglsres.cdmass.all$cor_av)
# Check equality of variances
leveneTest(cor_av ~ Adult_Habitat, data = pglsres.cdmass.all)

#SQUARE ROOT OF MASS
# Create a QQ plot 
qqnorm(pglsres.cdmass.all$rootmass_av)
qqline(pglsres.cdmass.all$rootmass)
# Histogram 
hist(pglsres.cdmass.all$rootmass, main = "Histogram of Root Mass")
# Check the normality 
shapiro.test(pglsres.cdmass.all$rootmass)
# Check equality of variances
leveneTest(rootmass_av ~ Adult_Habitat, data = pglsres.cdmass.all)

#RESIDUALS OF CORNEA SIZE~ROOTMASS
# Create a QQ plot 
qqnorm(pglsres.cdmass.all$pglsres.cdmass)
qqline(pglsres.cdmass.all$pglsres.cdmass)
# Histogram 
hist(pglsres.cdmass.all$pglsres.cdmass, main = "Histogram of cornea ~ mass residuals")
# Check the normality 
shapiro.test(pglsres.cdmass.all$pglsres.cdmass)
# Check equality of variances
leveneTest(pglsres.cdmass ~ Adult_Habitat, data = pglsres.cdmass.all)
```

Data for eye size and investment across ecological groups were not normally distributed, so we used non-parametric statistical tests for comparing data between ecological groups. 

?We used Kruskal-Wallis tests to test for differences in means across ecological groups, which is a non-parametric alternative to a one-way ANOVA test that uses ranks. 

When the Kruskal-Wallis test indicated that means differed across ecological groups and there were more than two groups, we used post-hoc Pairwise Wilcoxen Rank Sum Tests to calculate pairwise comparisons between group levels with corrections for multiple testing. This is a non-parametric alternative to a Tukey test that uses ranks to test for differences in means.

# Habitat PGLS

```{r}
# Run habitat pgls--------

pgls_hab1 <- pgls(cor_av ~ Adult_Habitat,
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#main effects
anova(pgls_hab1)

#print model output 
summary(pgls_hab1)

#PGLS cornea diameter vs. cube root of mass + habitat
pgls_hab <- pgls(log10(cor_av) ~ log10(rootmass_av) + Adult_Habitat,
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_hab)
par(mfrow = c(1, 1))

#Likelihood plot for Pagel's lambda. Solid red line indicates estimate for lambda and broken red lines indcaite the 95% confidence interval
lambda.hab <- pgls.profile(pgls_hab, "lambda")
plot(lambda.hab)

#main effects
anova(pgls_hab)

#print model output 
summary(pgls_hab)
```

## PGLS of eye scaling with body mass with habitat as covariate
```{r}
#run PGLS model for cornea diameter vs. cube root of mass using the Maximum Liklihood estimate of lambda
pgls_hab2 <- pgls(log10(cor_av) ~ log10(rootmass_av) * Adult_Habitat, 
               data = sal.comp, 
               lambda = "ML", bounds = list(lambda = c(1e-05, 1)),
               #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#evaluate model assumptions
par(mfrow = c(2,2)) #makes your plot window into 2x2 panels
plot(pgls_hab2) #plot the linear model
par(mfrow = c(1,1)) #set plotting window back to one plot

#significance of relationships
kable(anova(pgls_hab2), digits = 3, caption= "ANOVA Table for cornea size ~ mass * habitat") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1, valign = "top")

#intercept and slope
summary(pgls_hab2)
```
# Linear regression for cornea investment vs. adult habitat
```{r}

# Change Adult_Habitat to a factor with levels
pglsres.cdmass.all$Adult_Habitat <- factor(pglsres.cdmass.all$Adult_Habitat, levels = c("aquatic", "subfossorial", "terrestrial", "semiaquatic", "scansorial")) 

# Fit lineaar regression model for (absolute) cornea size and adult habitat
habitat.regression <- lm(cor_av ~ Adult_Habitat, data = pglsres.cdmass.all)
summary(habitat.regression)
anova(habitat.regression)

# Tukey's HSD test for pairwise comparisons betweens levels of Adult Habitat
tukey_abs <- glht(habitat.regression, linfct = mcp(Adult_Habitat = "Tukey"))
summary(tukey_abs)

# Fit lineaar regression model for relative cornea size and adult habitat
habitat.regression.1 <- lm(pglsres.cdmass ~ Adult_Habitat, data = pglsres.cdmass.all)
summary(habitat.regression.1)
anova(habitat.regression.1)

# Tukey's HSD test for pairwise comparisons between levels of Adult Habitat
tukey_inv <- glht(habitat.regression, linfct = mcp(Adult_Habitat = "Tukey"))
```

# Kruskal Wallis Test for cornea investment vs. adult habitat 

```{r}
# Test for differences in mean (absolute) cornea size across habitat groups: Kruskal-Wallis test
kruskal.test(cor_av ~ Adult_Habitat, data = pglsres.cdmass.all) 

# Multiple comparisons for differences in mean (absolute) cornea size across habitat groups: Pairwise Wilcox rank sum tests
pairwise.wilcox.test(pglsres.cdmass.all$cor_av, pglsres.cdmass.all$Adult_Habitat, p.adjust.method = "bonferroni")

# Multiple comparisons for differences in mean (absolute) cornea size across habitat groups: Dunn's post hoc test
dunn.test.abs <- dunn.test(pglsres.cdmass.all$cor_av, pglsres.cdmass.all$Adult_Habitat, method = "bonferroni")

# Test for differences relative cornea size (by mass) across habitat groups: Kruskal-Wallis test
kruskal.test(pglsres.cdmass ~ Adult_Habitat, data = pglsres.cdmass.all) 

# Multiple comparisons for differences in cornea investment (by mass) across habitat groups: Pairwise Wilcox rank sum tests
pairwise.wilcox.test(pglsres.cdmass.all$pglsres.cdmass, pglsres.cdmass.all$Adult_Habitat, p.adjust.method = "bonferroni")

# Multiple comparisons for differences in mean (absolute) cornea size across habitat groups: Dunn's post hoc test
dunn.test.inv <- dunn.test(pglsres.cdmass.all$pglsres.cdmass, pglsres.cdmass.all$Adult_Habitat, method = "bonferroni")
```

# Linear Regression and ANOVA

```{r}
# linear regression model
lm_adulthabitat <- lm(cor_av~ Adult_Habitat + rootmass_av, data = pglsres.cdmass.all)
summary(lm_adulthabitat)
anova(lm_adulthabitat)
```

# Activity Period PGLS

```{r}
# Run activity period pgls--------

pgls_act1 <- pgls(cor_av ~ Activity_Period,
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#main effects
anova(pgls_act1)

#print model output 
summary(pgls_act1)

#PGLS cornea diameter vs. cube root of mass + habitat
pgls_act <- pgls(log10(cor_av) ~ log10(rootmass_av) + Activity_Period,
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_act)
par(mfrow = c(1, 1))

#Likelihood plot for Pagel's lambda. Solid red line indicates estimate for lambda and broken red lines indcaite the 95% confidence interval
lambda.act <- pgls.profile(pgls_act, "lambda")
plot(lambda.act)

#main effects
anova(pgls_act)

#print model output 
summary(pgls_act)
```

# Larval Habitat PGLS

```{r}
# Run larval habitat pgls--------

pgls_larv1 <- pgls(cor_av ~ Larval_Habitat,
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#main effects
anova(pgls_larv1)

#print model output 
summary(pgls_larv1)

#PGLS cornea diameter vs. cube root of mass + habitat
pgls_larv <- pgls(log10(cor_av) ~ log10(rootmass_av) + Larval_Habitat,
               data = sal.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_larv)
par(mfrow = c(1, 1))

#Likelihood plot for Pagel's lambda. Solid red line indicates estimate for lambda and broken red lines indcaite the 95% confidence interval
lambda.larv <- pgls.profile(pgls_larv, "lambda")
plot(lambda.larv)

#main effects
anova(pgls_larv)

#print model output 
summary(pgls_larv)
```

# Aquatic PGLS

```{r}
# Run aquatic pgls--------

pgls_aq1 <- pgls(cor_av ~ Development,
               data = aquatic.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#main effects
anova(pgls_aq1)

#print model output 
summary(pgls_aq1)

#PGLS cornea diameter vs. cube root of mass + habitat
pgls_aq <- pgls(log10(cor_av) ~ log10(rootmass_av) + Development,
               data = aquatic.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_aq)
par(mfrow = c(1, 1))

#Likelihood plot for Pagel's lambda. Solid red line indicates estimate for lambda and broken red lines indcaite the 95% confidence interval
lambda.aq <- pgls.profile(pgls_aq, "lambda")
plot(lambda.aq)

#main effects
anova(pgls_aq)

#print model output 
summary(pgls_aq)
```