---
title: "Comparison of salamander eye size with other vertebrate groups"
author: "Katie Thomas & Nicole Bednarik"
date: 10 June 2024
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: show
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(smatr)
library(picante)
library(nlme)
library(ape)
library(evobiR)
library(geiger)
library(caper)
library(phytools)
library(phylotools)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
library(ggtree)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(mvMORPH)
library(tidyverse)

# Markdown settings
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)

```

# Import data and phylogenies

## Frogs

```{r load-data}

# Morphology data 

# Frog data (From Thomas et al. 2020 Proc B)
frog_morph <- data.frame(read.csv("../Data/Thomas et al. 2020/Data/frogs_analysis.csv", header=TRUE, na.strings=c("", "NA", " ")))

# Tidy data
frogs <- frog_morph %>%
  mutate(rootmass = Mass_g^(1/3)) %>%  
  mutate(eyemean = rowMeans(frog_morph[c('ED_right_mm', 'ED_left_mm')], na.rm=TRUE)) %>% 
  mutate(cormean = rowMeans(frog_morph[c('CD_right_mm', 'CD_left_mm')], na.rm=TRUE)) %>%  
  mutate(eyemean = na_if(eyemean, NaN)) %>% 
  mutate(cormean = na_if(cormean, NaN)) %>% 
  select(Order, genus_species, mod_tiplabel, SVL_mm, rootmass, eyemean, cormean) 

# Frog tree (from Jetz & Pyron, as modified in Thomas et al. 2020 Proc B)

# Read tree
frog.tree <- read.nexus(file = "../Data/Thomas et al. 2020/Data/frog-tree.nex")

#check that tree is dichotomous
is.binary(frog.tree)

#plot tree
plot.phylo(frog.tree, show.tip.label = FALSE)
```

## Salamanders

```{r load-salamanders}

# Import tidied salamander eye size and trait data
av.sal <- data.frame(read.csv("../Data/Tidy/salamanders_averages.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))

# Import pruned phylogeny
caudatatree <- read.nexus("../Data/Tidy/caudata-tree.nex")
```

## Fishes
```{r load-fishes}

# Load vertebrate eye size data assembled from literature
vert_morph <- data.frame(read.csv("../Data/Thomas et al. 2020/Data/other verts/vertebrate_eyesize.csv", header=TRUE, na.strings=c("", "NA", " ")))

# Tidy data
vert_morph <- vert_morph %>%
  mutate(rootmass = (Body_mass_kg*1000)^(1/3)) %>%
  filter(!(is.na(VertLife_name)&is.na(TimeTree_tiplabel)&is.na(TimeTree_genus_sub))) %>%
  mutate_if(is.character, as.factor)

# Pull out fishes
fishes <- vert_morph[vert_morph$Group == "Fishes (Actinopterygii)", ]

# Species means for fishes
fishes_avs <- fishes %>% 
  mutate(tips = coalesce(TimeTree_tiplabel, TimeTree_genus_sub)) %>% 
  mutate_if(is.character, as.factor) %>% 
  group_by(tips) %>%
  summarise(TD_av = mean(Trans_diameter_mm),
            mass_av = mean(rootmass), 
            svl_av = mean(SVL_mm),
            n = n())

# Fish phylogeny
fishes.phy <- read.newick(file = "../Data/Thomas et al. 2020/Data/other verts/vertebrate_trees/TimeTree_fishes.nwk") 

# Check that tree is dichotomous
is.binary(fishes.phy)

# Plot tree
plot.phylo(fishes.phy, show.tip.label = FALSE)

```

## Squamates
```{r load-squamates}

# Squamate data
squamates <- vert_morph[vert_morph$Group == "Reptiles (Squamata)", ]

# Species means
squamate_avs <- squamates %>% 
  mutate(VertLife_name = gsub('\\s+', '_', VertLife_name)) %>%
  mutate_if(is.character, as.factor) %>% 
  group_by(VertLife_name, Subgroup) %>%
  summarise(TD_av = mean(Trans_diameter_mm),
            CD_av = mean(Cornea_diameter_mm),
            mass_av = mean(rootmass), 
            svl_av = mean(SVL_mm),
            n = n()) %>%
  droplevels()

# Squamate phylogeny

# Read tree
squamate.phy <- read.tree(file = "../Data/Thomas et al. 2020/Data/other verts/vertebrate_trees/VertLifesquamate_tree/Vert_Life_Squamate_consensus_tree.newick")
  
#check that tree is dichotomous
is.binary(squamate.phy)

#make dichotomous (randomly resolve polytomies)
squamate.phy <- multi2di(squamate.phy)

#plot tree
plot.phylo(squamate.phy, show.tip.label = FALSE)
```



# Cornea diameter vs. mass

## Frogs

```{r frog-means}

# Find species means for cornea diameter vs. rootmass
frogs.cdmass <- frogs %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(cormean) | is.na(rootmass))) %>% 
  group_by(genus_species, mod_tiplabel) %>%
  summarise(cor_av = mean(cormean), mass_av = mean(rootmass), n = n()) %>%
  mutate_if(is.factor, as.character)

frogs.cdmass <- as.data.frame(frogs.cdmass)
```

Match data to tree & fit PGLS in caper

```{r frog-PGLS}

# Make row names of dataset the species names (so it will match phylogeny tips)
rownames(frogs.cdmass) <- frogs.cdmass$mod_tiplabel

# Check that names match in dataframe and tree
name.check(phy = frog.tree, data = frogs.cdmass)

# Use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
frog.cdmass.comp <- comparative.data(phy = frog.tree, data = frogs.cdmass, 
                            names.col = mod_tiplabel, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

# Check for dropped tips or dropped species
frog.cdmass.comp$dropped$tips #phylogeny
frog.cdmass.comp$dropped$unmatched.rows #dataset

# Run PGLS model
pgls_frog.cdmass <- pgls(log10(cor_av) ~ log10(mass_av), 
               data = frog.cdmass.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

# Check model assumptions with diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_frog.cdmass)

summary(pgls_frog.cdmass)

```

## Salamanders 

```{r salamanders-PGLS}

# Make row names of dataset the species names (so it will match phylogeny tips)
rownames(av.sal) <- av.sal$Genus_Species

# Check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.sal)

# Use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp <- comparative.data(phy = caudatatree, data = av.sal, 
                            names.col = Genus_Species, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

# Check for dropped tips or dropped species
sal.comp$dropped$tips #phylogeny
sal.comp$dropped$unmatched.rows #dataset

# Run PGLS model
pgls_salamanders.cdmass <- pgls(log10(cor_av) ~ log10(rootmass_av), 
               data = sal.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

# Check model assumptions with diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_salamanders.cdmass)

summary(pgls_salamanders.cdmass)
```

## Plot for cornea vs. mass

```{r}

# Color vector
col_cdmass <- c("frogs" = "#009E73",
            "salamanders" = "#9900ff")

shape_cdmass <- c("frogs" = 19,
                  "salamanders" = 15)

#plot data
plot_cdmass <- ggplot(frogs.cdmass, aes(x = mass_av, y = cor_av, color = "frogs", pch = "frogs")) +
  geom_point(alpha = 0.5, size = 2.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.text.align = 0, 
        axis.text = element_text(size = 17), 
        axis.title = element_text(size = 17), 
        axis.ticks = element_line(size = 0.8),
        axis.ticks.length = unit(0.15, "cm"),
        panel.border = element_rect(size = 1.45),
        axis.text.x = element_text(margin = margin(t = 7)),
        axis.text.y = element_text(margin = margin(r = 7))) +
  scale_y_log10(name = "cornea diameter (mm)") +
  scale_x_log10(name = "cube root of mass (g)") +
  geom_abline(slope = coef(pgls_frog.cdmass)[[2]], intercept = coef(pgls_frog.cdmass)[[1]], color = col_cdmass["frogs"], size = 1.15) +
  #add salamanders to plot
  geom_point(data = pglsres.cdmass.all, aes(x = rootmass_av, y = cor_av, color = "salamanders", pch = "salamanders"), alpha = 0.5, size = 2.5) +
  geom_abline(slope = coef(pgls_salamanders.cdmass)[[2]], intercept = coef(pgls_salamanders.cdmass)[[1]], color = col_cdmass["salamanders"], size = 1.15) +
  scale_color_manual(values = col_cdmass, name = "Group") +
  scale_shape_manual(values = shape_cdmass, name = "Group")

#show plot
plot_cdmass
```

# Eye diameter vs. mass

## Frogs

```{r}

# Find species means for eye diameter vs. rootmass
frogs.edmass <- frogs %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(eyemean) | is.na(rootmass))) %>% 
  group_by(genus_species, mod_tiplabel) %>%
  summarise(eye_av = mean(eyemean), mass_av = mean(rootmass), n = n()) %>%
  mutate_if(is.factor, as.character)

frogs.edmass <- as.data.frame(frogs.edmass)
```

Match data to tree & fit PGLS in caper

```{r}

# Make row names of dataset the species names (so it will match phylogeny tips)
rownames(frogs.edmass) <- frogs.edmass$mod_tiplabel

# Check that names match in dataframe and tree
name.check(phy = frog.tree, data = frogs.edmass)

# Use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
frog.edmass.comp <- comparative.data(phy = frog.tree, data = frogs.edmass, 
                            names.col = mod_tiplabel, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

# Check for dropped tips or dropped species
frog.edmass.comp$dropped$tips #phylogeny
frog.edmass.comp$dropped$unmatched.rows #dataset

# Run PGLS model
pgls_frog.edmass <- pgls(log10(eye_av) ~ log10(mass_av), 
               data = frog.edmass.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

# Check model assumptions with diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_frog.edmass)

summary(pgls_frog.edmass)
```


## Salamanders

```{r}

# Run PGLS model
pgls_sal.edmass <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = sal.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#check model assumptions with diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_sal.edmass)

summary(pgls_sal.edmass)
```

## Fishes

Match data to tree & fit PGLS in caper

```{r}

# Make row names of dataset the species names (so it will match phylogeny tips)
fishes_avs <- as.data.frame(fishes_avs)
rownames(fishes_avs) <- fishes_avs$tips

# Check that names match in dataframe and tree
name.check(phy = fishes.phy, data = fishes_avs)

# Use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
fishes.comp <- comparative.data(phy = fishes.phy, data = fishes_avs, 
                            names.col = tips, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

# Check for dropped tips or dropped species
fishes.comp$dropped$tips #phylogeny
fishes.comp$dropped$unmatched.rows #dataset

# Run PGLS model
pgls_fish.edmass <- pgls(log10(TD_av) ~ log10(mass_av), 
               data = fishes.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

# Check model assumptions with diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_fish.edmass)

summary(pgls_fish.edmass)
```
## Plot for eye vs. mass

Note that the salamander eye size dataset is missing the species with the smallest eyes (those had cornea measurements only), so that pushes the regression line up quite a bit. 

```{r}

# Color vector
col_ed <- c("frogs" = "#009E73",
            "salamanders" = "#9900ff",
            "fishes" = "#0066ff")

shape_ed <- c("frogs" = 19,
            "salamanders" = 15,
            "fishes" = 17)

# Plot data
plot_edmass <- ggplot(frogs.edmass, aes(x = mass_av, y = eye_av, color = "frogs", pch = "frogs")) +
  geom_point(alpha = 0.5, size = 2.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.text.align = 0,
        axis.text = element_text(size = 17), 
        axis.title = element_text(size = 17), 
        axis.ticks = element_line(size = 0.8),
        axis.ticks.length = unit(0.15, "cm"),
        panel.border = element_rect(size = 1.45),
        axis.text.x = element_text(margin = margin(t = 7)),
        axis.text.y = element_text(margin = margin(r = 7))) +
  scale_y_log10(name = "eye diameter (mm)") +
  scale_x_log10(name = "cube root of mass (g)", breaks = c(1, 3, 10), limits=c(NA,10.01)) +
  geom_abline(slope = coef(pgls_frog.edmass)[[2]], intercept = coef(pgls_frog.edmass)[[1]], color = col_ed["frogs"], size = 1.15) +
  #add salamanders to plot
  geom_point(data = av.sal, aes(x = rootmass_av, y = eye_av, color = "salamanders", pch = "salamanders"), alpha = 0.5, size = 2.5) +
  geom_abline(slope = coef(pgls_sal.edmass)[[2]], intercept = coef(pgls_sal.edmass)[[1]], color = col_ed["salamanders"], size = 1.15) +
  #add fishes to plot
  geom_point(data = fishes_avs, aes(x = mass_av, y = TD_av, color = "fishes", pch = "fishes"), alpha = 0.5, size = 2.5) +
  geom_abline(slope = coef(pgls_fish.edmass)[[2]], intercept = coef(pgls_fish.edmass)[[1]], color = col_ed["fishes"], size = 1.15) +
  scale_color_manual(values = col_ed, name = "Group") +
  scale_shape_manual(values = shape_ed, name = "Group") 

#show plot
plot_edmass
```


# Eye diameter vs. SVL

## Frogs

```{r}

# Find species means for eye diameter vs svl
frogs.edsvl <- frogs %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(eyemean) | is.na(SVL_mm))) %>% 
  group_by(genus_species, mod_tiplabel) %>%
  summarise(eye_av = mean(eyemean), svl_av = mean(SVL_mm), n = n()) %>%
  mutate_if(is.factor, as.character)

frogs.edsvl <- as.data.frame(frogs.edsvl)
```

Match data to tree & fit PGLS in caper

```{r}

#Make row names of dataset the species names (so it will match phylogeny tips)
rownames(frogs.edsvl) <- frogs.edsvl$mod_tiplabel

#check that names match in dataframe and tree
name.check(phy = frog.tree, data = frogs.edsvl)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
frog.edsvl.comp <- comparative.data(phy = frog.tree, data = frogs.edsvl, 
                            names.col = mod_tiplabel, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
frog.edsvl.comp$dropped$tips #phylogeny
frog.edsvl.comp$dropped$unmatched.rows #dataset

#run PGLS model
pgls_frog.edsvl <- pgls(log10(eye_av) ~ log10(svl_av), 
               data = frog.edsvl.comp, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#check model assumptions with diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_frog.edsvl)

summary(pgls_frog.edsvl)

# Residuals for PGLS of eye diameter ~ SVL

# Extract PGLS residuals 
pglsres.edsvl.frog <- residuals(pgls_frog.edsvl) 

# Name residuals
colnames(pglsres.edsvl.frog) <- "pglsres.edsvl.frog" 

# Making PGLS vector into a dataframe
pglsres.edsvl.frog <- as.data.frame(pglsres.edsvl.frog)

# Exponentiate PGLS residuals for plots
pglsres.edsvl.frog <- pglsres.edsvl.frog %>%
  mutate(pglsres.edsvl.frog = 10^pglsres.edsvl.frog)
```




## Salamanders

```{r}

# Run PGLS model
pgls_sal.edsvl <- pgls(log10(eye_av) ~ log10(svl_av), 
               data = sal.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

# Check model assumptions with diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_sal.edsvl)

summary(pgls_sal.edsvl)

# Residuals for PGLS of eye diameter ~ SVL

# Extract PGLS residuals 
pglsres.edsvl.sal <- residuals(pgls_sal.edsvl) 

# Name residuals
colnames(pglsres.edsvl.sal) <- "pglsres.edsvl.sal" 

# Making PGLS vector into a dataframe
pglsres.edsvl.sal <- as.data.frame(pglsres.edsvl.sal)

# Exponentiate PGLS residuals for plots
pglsres.edsvl.sal <- pglsres.edsvl.sal %>%
  mutate(pglsres.edsvl.sal = 10^pglsres.edsvl.sal)
```

## Squamates

```{r}

# Subset groups with eye size data
squamate.ed <- squamate_avs %>%
  filter(Subgroup=="colubrid snakes" | Subgroup=="geckos") %>%
  droplevels()
squamate.ed <- as.data.frame(squamate.ed)
           
# Make row names of dataset the species names (so it will match phylogeny tips)
rownames(squamate.ed) <- squamate.ed$VertLife_name

# Check that names match in dataframe and tree
name.check(phy = squamate.phy, data = squamate.ed)

# Use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
squam.edsvl.comp <- comparative.data(phy = squamate.phy, data = squamate.ed, 
                            names.col = VertLife_name, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

# Check for dropped tips or dropped species
squam.edsvl.comp$dropped$tips #phylogeny
squam.edsvl.comp$dropped$unmatched.rows #dataset

# Run PGLS model
pgls_squam.edsvl <- pgls(log10(TD_av) ~ log10(svl_av) * Subgroup, 
               data = squam.edsvl.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

# Check model assumptions with diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_squam.edsvl)

summary(pgls_squam.edsvl)

# Residuals for PGLS of eye diameter ~ SVL

# Extract PGLS residuals 
pglsres.edsvl.squam <- residuals(pgls_squam.edsvl) 

# Name residuals
colnames(pglsres.edsvl.squam) <- "pglsres.edsvl.squam" 

# Making PGLS vector into a dataframe
pglsres.edsvl.squam <- as.data.frame(pglsres.edsvl.squam)

# Exponentiate PGLS residuals for plots
pglsres.edsvl.squam <- pglsres.edsvl.squam %>%
  mutate(pglsres.edsvl.squam = 10^pglsres.edsvl.squam)
```

## Plot for eye vs. SVL

Note that the salamander eye size dataset is missing the species with the smallest eyes (those had cornea measurements only), so that pushes the regression line up quite a bit. 

```{r}

# Color vector
col_edsvl <- c("frogs" = "#009E73",
            "salamanders" = "#9900ff",
            "geckos" = "#FF9933",
            "colubrid snakes" = "#FF3333")

shape_edsvl <- c("frogs" = 19,
            "salamanders" = 15,
            "geckos" = 17,
            "colubrid snakes" = 18)

# Plot data
plot_edsvl <- ggplot(frogs.edsvl, aes(x = svl_av, y = eye_av, color = "frogs", pch = "frogs")) +
  geom_point(alpha = 0.5, size = 2.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        legend.text.align = 0,
        axis.text = element_text(size = 17), 
        axis.title = element_text(size = 17), 
        axis.ticks = element_line(size = 0.8),
        axis.ticks.length = unit(0.15, "cm"),
        panel.border = element_rect(size = 1.45),
        axis.text.x = element_text(margin = margin(t = 7)),
        axis.text.y = element_text(margin = margin(r = 7))) +
  scale_y_log10(name = "eye diameter (mm)") +
  scale_x_log10(name = "snout-vent length (mm)") +
  geom_abline(slope = coef(pgls_frog.edsvl)[[2]], intercept = coef(pgls_frog.edsvl)[[1]], color = col_edsvl["frogs"], size = 1.15) +
  #add salamanders to plot
  geom_point(data = av.sal, aes(x = svl_av, y = eye_av, color = "salamanders", pch = "salamanders"), alpha = 0.5, size = 2.5) +
  geom_abline(slope = coef(pgls_sal.edsvl)[[2]], intercept = coef(pgls_sal.edsvl)[[1]], color = col_edsvl["salamanders"], size = 1.15) +
  #add squamates to plot
  geom_point(data = squamate.ed, aes(x = svl_av, y = TD_av, color = Subgroup, pch = Subgroup), alpha = 0.5, size = 2.5) +
  #snake line
  geom_abline(slope = coef(pgls_squam.edsvl)[[2]], intercept = coef(pgls_squam.edsvl)[[1]], color = col_edsvl["colubrid snakes"], size = 1.15) +
  #gecko line
  geom_abline(slope = coef(pgls_squam.edsvl)[[2]] + coef(pgls_squam.edsvl)[[4]], 
              intercept = coef(pgls_squam.edsvl)[[1]] + coef(pgls_squam.edsvl)[[3]], color = col_edsvl["geckos"], size = 1.15) +
  scale_color_manual(values = col_edsvl, name = "Group") +
  scale_shape_manual(values = shape_edsvl, name = "Group")
 
# Show plot
plot_edsvl
interactive_plot_edsvl <- ggplotly(plot_edsvl)
interactive_plot_edsvl
```

# Figure

We can put these plots together to make a paneled figure

```{r}

# Make panels 
fig.a <- plot_cdmass +
  theme(text = element_text(size=10))

fig.b <- plot_edmass +
  theme(text = element_text(size=10))

fig.c <- plot_edsvl +
  theme(text = element_text(size=10))



# Use cowplot package to nicely plot the graphs together with a shared common legend.
plots <- plot_grid(fig.a, fig.b, fig.c, 
                   align = 'vh', 
                   hjust = -1, 
                   vjust = 1,
                   ncol = 2,
                   rel_widths = c(1,1))

# Show figure
plots

# Export pdf of figure
pdf("../Plots/vert-comparison.pdf", width = 10, height = 5)
plot_grid(plots)
dev.off()
```

